### 2023-09-10 chapters/background.tex
First complete introduction

The \emph{information entropy} of a random variable $X$ is defined as the \emph{expected information content} of an outcome of $X$,
%
\begin{equation} \label{eq:information-entropy}
    H(X) = \mathbb{E}_X\left[I_X(x)\right] = - \sum_{x\in X} p_X(x) \log p_X(x) \enspace .
\end{equation}
%
This means that $H(X)$ measures the amount of information we can expect to gain from observing an outcome of $X$, when we know only its distribution $p_X(x)$. If all possible outcomes are equally likely, then $H(X)$ is maximized, and if only one outcome is possible, then $H(X)$ is minimized.

@@

To generalize the concept of information entropy to continuous random variables, Shannon originally replaced the sum over the probability mass function in \eqref{eq:information-entropy} with an integral over the probability density function, as suggested by the definition $H(X) = \mathbb{E}_X\left[I_X(x)\right]$. This approach leads to the definition of the information entropy of a continuous random variable $X$ as, 
%
\begin{equation} \label{eq:differential-entropy}
    H(X) = - \int_{\mathcal{X}} p_X(x) \log p_X(x) \, dx \enspace ,
\end{equation}
%
which is called the \emph{differential entropy} \cite{shannon_mathematical_1948}. Here, $\mathcal{X}$ is the support of $X$.

@@

Instead, \textcite{jaynes_information_1957} argued that the information entropy of a continuous distribution should be defined as the limiting density of increasingly dense discrete distributions \cite{jaynes_prior_1968}. This argument leads to an expression for the information entropy of a continuous random variable $X$ as,
%
\begin{equation} \label{eq:corrected-differential-entropy-jaynes}
    H_{N\rightarrow\infty}(X) = \lim_{N\rightarrow\infty} [H_{N}(X) - \log N] = - \int_\mathcal{X} p(x) \log \left(\frac{p(x)}{q(x)}\right) \enspace ,
\end{equation}
%
where $H_N(X)$ is the limiting density for discrete points, $q(x)$ is a uniform density over the quantization of the continuous space, and we have subtracted a $\log N$ term that would go to infinity in the limit of infinite points.
By doing this, the information entropy becomes dimensionally correct, and invariant under a change of variables%
\footnote{\label{fn:change-of-variables-invariance-of-corrected-differential-entropy}
    Under a change of independent variable from $x$ to $y(x)$, we have that
    \begin{align*}
        \widetilde{w}(y)\,dy = w(x)\,dx \enspace , \quad
        \widetilde{q}(y)\,dy = q(x)\,dx \enspace .
    \end{align*}
    Plugging this into the differential entropy \cref{eq:corrected-differential-entropy-jaynes} we arrive at the original expression but now with $y$ as the independent variable,
    \begin{equation*}
        H_{N\rightarrow\infty}(y(x)) =  = - \int_\mathcal{Y} \widetilde{w}(y)\,\frac{dy}{dx} \log \frac{\widetilde{w}(y)\frac{dy}{dx}}{\widetilde{q}(y)\frac{dy}{dx}} \, dx = - \int_\mathcal{Y} w(y) \log \frac{w(y)}{q(y)} \, dy \enspace .
    \end{equation*}
}.

The form on the right-hand side of \cref{eq:corrected-differential-entropy-jaynes} can be recognized as the negative Kullback-Leibler divergence of $p(x)$ to $q(x)$ \cite{kullback_information_1959},
%
\begin{equation} \label{eq:kullback-leibler-divergence}
    D_{KL}(P\parallel Q) = \int_\mathcal{X} p(x) \log \left(\frac{p(x)}{q(x)}\right) \enspace .
\end{equation}
%
The Kullback-Leibler divergence $D_{KL}(p\parallel q)$ is often interpreted as the amount of additional information required to represent events from distribution the $P$ with density $p(x)$ using a code optimized for distribution $Q$ with density $q(x)$. In other words, it measures how surprised one would be if they used distribution $Q$ to represent events from distribution $P$.
The KL-divergence is also known as the \emph{relative entropy} of $P$ with respect to $Q$, or simply as the \emph{information gain}.

Returning to Jaynes and \cref{eq:corrected-differential-entropy-jaynes}, the argument can be interpreted to say that the information entropy of a continuous random variable $X$ should be defined in relative terms as the expected difference in information entropy between its actual density and a uniform density.
The Kullback-Leibler divergence as a measure of (relative) information entropy also provides a natural point of convergence of this section since it is equivalently defined for both discrete and continuous random variables.

@@

\subsection{Two kinds of uncertainty}

\todo[inline]{Consider not using the terms aleatoric and epistemic uncertainty. Instead use the terms ``uncertainty due to lack of knowledge (reducible)" and ``uncertainty inherent to the data (irreducible)".}

Uncertainty can be divided into two kinds: aleatoric and epistemic.

Aleatoric uncertainty is defined as being inherent to the data, and is irreducible.

Epistemic uncertainty is uncertainty due to lack of knowledge, and can be reduced with more data.

>>>>

The \emph{information entropy of a discrete random variable} $X$ is defined as the \emph{expected information content} of an outcome of $X$,
%
\begin{equation} \label{eq:information-entropy}
    H(X) = \mathbb{E}_X\left[I_X(x)\right] = - \sum_{x\in X} p_X(x) \log p_X(x) \enspace .
\end{equation}
%
This means that $H(X)$ measures the amount of information we can expect to gain from observing an outcome of $X$, when we know only its distribution $p_X(x)$. If all possible outcomes are equally likely, then $H(X)$ is maximized, and if only one outcome is possible, then $H(X)$ is minimized.

@@

To generalize the concept of information entropy to continuous random variables, Shannon originally replaced the sum over the probability mass function in \eqref{eq:information-entropy} with an integral over the probability density function, as suggested by the definition $H(X) = \mathbb{E}_X\left[I_X(x)\right]$. This approach leads to the definition of the \emph{information entropy of a continuous random variable} $X$ as, 
%
\begin{equation} \label{eq:differential-entropy}
    H(X) = - \int_{\mathcal{X}} p_X(x) \log p_X(x) \, dx \enspace ,
\end{equation}
%
which is called the \emph{differential entropy} \cite{shannon_mathematical_1948}. Here, $\mathcal{X}$ is the support of $X$.

@@

Instead, \textcite{jaynes_information_1957,jaynes_prior_1968} argued that the information entropy of a continuous distribution should be defined as the limiting density of increasingly dense discrete distributions, $H_{N\rightarrow\infty}(X)$. This argument leads to,
%
\begin{equation} \label{eq:corrected-differential-entropy-jaynes}
    H_{N\rightarrow\infty}(X) = \lim_{N\rightarrow\infty} [H_{N}(X) - \log N] = - \int_\mathcal{X} p(x) \log \left(\frac{p(x)}{q(x)}\right) \enspace ,
\end{equation}
%
where $H_N(X)$ is the limiting density for discrete points, $q(x)$ is a uniform density over the quantization of the continuous space, and we have subtracted a $\log N$ term that would go to infinity in the limit of infinite points.
By doing this, the information entropy becomes positive, dimensionally correct, and invariant under a change of variables%
\footnote{\label{fn:change-of-variables-invariance-of-corrected-differential-entropy}
    Under a change of independent variable from $x$ to $y(x)$, we have that
    \begin{align*}
        \widetilde{w}(y)\,dy = w(x)\,dx \enspace , \quad
        \widetilde{q}(y)\,dy = q(x)\,dx \enspace .
    \end{align*}
    Plugging this into the differential entropy \cref{eq:corrected-differential-entropy-jaynes} we arrive at the original expression but now with $y$ as the independent variable,
    \begin{equation*}
        H_{N\rightarrow\infty}(y(x)) =  = - \int_\mathcal{Y} \widetilde{w}(y)\,\frac{dy}{dx} \log \frac{\widetilde{w}(y)\frac{dy}{dx}}{\widetilde{q}(y)\frac{dy}{dx}} \, dx = - \int_\mathcal{Y} w(y) \log \frac{w(y)}{q(y)} \, dy \enspace .
    \end{equation*}
}.

The form on the right-hand side of \cref{eq:corrected-differential-entropy-jaynes} can be recognized as the negative Kullback-Leibler divergence of $p(x)$ to $q(x)$ \cite{kullback_information_1959},
%
\begin{equation} \label{eq:kullback-leibler-divergence}
    D_{KL}(P\parallel Q) = \int_\mathcal{X} p(x) \log \left(\frac{p(x)}{q(x)}\right) \enspace .
\end{equation}
%
The Kullback-Leibler divergence $D_{KL}(P\parallel Q)$ is often interpreted as the amount of additional information required to represent events from distribution $P$ with density $p(x)$ using a code optimized for distribution $Q$ with density $q(x)$. In other words, it measures how surprised one would be if they used distribution $Q$ to represent events from distribution $P$, i.e. the \emph{relative entropy} of $P$ with respect to $Q$. 
The KL-divergence is also known as the \emph{information gain}.

Returning to Jaynes' argument and \cref{eq:corrected-differential-entropy-jaynes}, we can interpret this to say that the information entropy of a continuous random variable $X$ should be defined as the expected difference in information entropy between its density and a uniform density. 
Identifying the Kullback-Leibler divergence as a measure of (relative) information entropy provides a natural point of convergence for this section since it is equivalently defined for both discrete and continuous random variables as opposed to the differential entropy. 
In this view, information entropy (relative or absolute) is a quantity that distills the distribution of a random variable into a single number that describes the diversity of the potential outcomes of the random variable.

@@

% \subsection{Two kinds of uncertainty}

% \todo[inline]{Consider not using the terms aleatoric and epistemic uncertainty. Instead use the terms ``uncertainty due to lack of knowledge (reducible)" and ``uncertainty inherent to the data (irreducible)".}

% Uncertainty can be divided into two kinds: aleatoric and epistemic.

% Aleatoric uncertainty is defined as being inherent to the data, and is irreducible.

% Epistemic uncertainty is uncertainty due to lack of knowledge, and can be reduced with more data.
